# Red Wine Quality Analysis
### Which chemical properties influence the quality of red wines?

By Linjing Yang

# Introduction
The tidy data set contains 1,599 red wines [^1] with 11 variables on the chemical properties of the wine. At least 3 wine experts rated the quality of each wine, providing a rating between 0 (very bad) and 10 (very excellent).

+ 1599 records
+ 11 chemical properties
+ quality rating: 0 - 10

## Variables

###Input variables (based on physicochemical tests):
1. fixed acidity (tartaric acid - g / dm^3)
2. volatile acidity (acetic acid - g / dm^3)
3. citric acid (g / dm^3)
4. residual sugar (g / dm^3)
5. chlorides (sodium chloride - g / dm^3
6. free sulfur dioxide (mg / dm^3)
7. total sulfur dioxide (mg / dm^3)
8. density (g / cm^3)
9. pH
10. sulphates (potassium sulphate - g / dm3)
11. alcohol (% by volume)
   
###Output variable (based on sensory data): 
12. quality (score between 0 and 10)





# Exploration of Data
```{r echo=FALSE}
# {r echo=FALSE} Do not show this code block.

# Avoid all codes and warnings to be shown in the HTML document
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)
```

```{r}
### Load libraries
library(GGally)
library(ggplot2)
library(gridExtra)
library(RColorBrewer) # Scatter Plot: color
library(dplyr) # %>%
library(tidyr)
library(memisc) # Linear Regression: mtable
library(corrplot) # correlation matrix
library(ggridges) # Ridge Line Plot
library(kableExtra) # Show the head of the orginal data
library(tibble) # Check distribution of the whole group of chemicals
library(car) # Scatterplot Matrix
library(corrplot) # Correlation Matrix (with cor. number)

library(ellipse)
library(RColorBrewer)
library(ggpubr) # add cor coefficient to scatter plot
```

```{r results = 'hide'}
#{r results = 'hide'} Do not show the result of this code block.

### Load the data
getwd()
list.files()
df <- read.csv('wineQualityReds.csv')
head(df)
```

```{r results = 'hide'}
# Details of data
str(df)
```

```{r}
# show the head of the data

# Subset: drop 1st line - X
# tmpdata <- subset(df, select = -c(varname))
df_s <- subset(df, select = -c(X))

# Draw the table
df_s %>% head(6) %>% kable() %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed", "responsive"), 
    full_width = F
  )
```

```{r}
# All chemicals in df
chemicals <-  list(
                df$alcohol,
                df$citric.acid,
                df$chlorides,
                df$density,
                df$fixed.acidity, 
                df$free.sulfur.dioxide, 
                df$pH,
                df$residual.sugar,
                df$sulphates,
                df$volatile.acidity,
                df$total.sulfur.dioxide
                  )

# List:
#   Index of lists starts from [1] rather than [0]
#   Access the element of a list: list[[index]]

```

```{r}
# Names of all chemicals
chemical_names <- c(
    'Alcohol',
    'Citric Acid',
    'Chlorides',
    'Density',
    'Fixed Acidity', 
    'Free Sulfur Dioxide', 
    'pH Value',
    'Residual Sugar',
    'Sulphates',
    'Volatile Acidity',
    'Total Sulfur Dioxide'
)
```

```{r}
# Plot arguments of chemicals

# c(start, end, step)

chemical_units <- list(
                      #alcohol  1
                      '(% by volume)',
                      
                      #citric.acid 2
                      '(g / dm^3)',
                      
                      #chlorides 3
                      '(g / dm^3)',
                      
                      #density 4
                      '(g / cm^3)',
                      
                      #fixed.acidity 5
                      '(g / dm^3)', 
                      
                      #free.sulfur.dioxide 6 
                      '(mg / dm^3)', 
                      
                      #pH 7
                      '', 
                      
                      #residual.sugar 8 
                      '(g / dm^3)', 
                      
                      #sulphates 9
                      '(g / dm3)',
                      
                      #volatile.acidity 10
                      '(g / dm^3)', 
                      
                      #total.sulfur.dioxide 11
                      '(mg / dm^3)' 
                    )
```

```{r}
# Plot arguments of chemicals

# c(start, end, step)

chemical_args <- list(
                      #alcohol  1
                      c(8, 15, 1),
                      
                      #citric.acid 2
                      c(0, 1, 0.1),
                      
                      #chlorides 3
                      c(0, 0.2, 0.02),
                      
                      #density 4
                      c(0.99, 1.0037, 0.01),
                      
                      #fixed.acidity 5
                      c(4, 16, 1), 
                      
                      #free.sulfur.dioxide 6 
                      c(1, 72, 10), 
                      
                      #pH 7
                      c(2.7, 4.1, 0.5), 
                      
                      #residual.sugar 8 
                      c(1, 5, 0.5), 
                      
                      #sulphates 9
                      c(0.3, 1.3, 0.2),
                      
                      #volatile.acidity 10
                      c(0.1, 1.2, 0.2), 
                      
                      #total.sulfur.dioxide 11
                      c(6, 160, 50) 
                    )
```

```{r}
# Function: Density plot

create_density <- 
  function(chemical, chemical_name, start, end, step, chemical_unit){
    p <- ggplot(aes(x=chemical), 
                data =df) +
         geom_density(alpha = my_alpha, fill= my_color, color = my_color) +
         scale_x_continuous(limits = c(start, end),
                            breaks = seq(start, end, step))+
         labs(
           x = paste(chemical_name, chemical_unit),
           y = 'Density')
    return(p)
  }
```

```{r}
# Function: create 3 histograms: orginal, log10, sqrt
# And put density plot together

create_4_plots_table <- 
  function(chemical, chemical_name, start, end, step, chemical_unit){
    p1 <- ggplot(aes(x=chemical), data= df) +
            geom_histogram(alpha = my_alpha, fill= my_color) +
            xlab(paste(chemical_name, chemical_unit, sep = ' '))
    
    p2 <- p1 + scale_x_log10() + 
               xlab(paste(chemical_name, "(log10)", sep = ' '))
    
    p3 <- p1 + scale_x_sqrt() + 
               xlab(paste(chemical_name, "(sqrt)", sep = ' '))
  
    p0 <- create_density(chemical, chemical_name, 
                         start, end, step, chemical_unit)
    
    p <- grid.arrange(p0, p1, p2, p3, 
                      ncol=2, 
                      top = chemical_name # Name of the four plots
                      ) 
    return(p)
  }
```

```{r}
# set color arguments

my_color <- '#008AD8'   # Main Color of this report: blue

# "#BB4444", "#EE9988" ,'#f9e0db', '#008AD8', #68c6f9,'#cfeaf9'
red <- c("#BB4444", "#EE9988" ,'#f9e0db')

blue <- c('#0167a3','#008AD8', '#68c6f9','#92cdef','#cfeaf9')

my_alpha <- 0.7
```

```{r}
# Set themes

my_theme <- theme(
                plot.title = element_text(size = 20),  
                axis.title.x = element_text(size = 14),  
                axis.title.y = element_text(size = 14),
                
                legend.title = element_text(size = 14),
                legend.text = element_text(size = 12),
                axis.text.x = element_text(angle = 90, hjust = 1) #Rotate axis labels
             )
```


```{r}
# Scatter plot: Lines
line_mean <- geom_line(stat='summary', fun.y=mean, 
                       linetype=1, color='red')

line_median <- geom_line(stat='summary', fun.y=quantile,
                         fun.args=list(probs=.5),
                         linetype=1, color=my_color)
```

```{r}
# Scatter plot: Regression Line
line_regression <- geom_smooth(method='lm',formula=y~x, 
                               color=my_color, fill = "skyblue"
                               )
```


```{r}
# scatter plot:  Add correlation coefficient (r)

# library(ggpubr)
text_cor <- stat_cor(method = "spearman")
```


```{r}
# Scatter Plot: Quality v.s. Chemical

# Draw Plot
create_scatter <- 
  function(chemical, chemical_name, start, end, step, chemical_unit){
    p <- ggplot(aes(x=chemical, y=quality),data=df)+
          geom_point(alpha = 0.1, size = 1, 
                     position = 'jitter', color=my_color) +
          scale_x_continuous(limits = c(start, end),
                            breaks = seq(start, end, step)) +
          labs(
              title = paste(chemical_name, 'and Quality'),
              x = paste(chemical_name, chemical_unit),
              y = "Quality"
           ) +
          line_regression +
          line_mean +
          text_cor
    return(p)
  }
```



```{r}
#test

#create_scatter(df$alcohol, 'Alcohol', 8, 15, 1, 'mg' )
```



```{r results = 'hide'}
#Ridge Line Plot

# Step 1: group df by quality

df$quality.group <-
  cut(df$quality,
      c(3,4,5,6,8))

table(df$quality.group,
      useNA = 'ifany')
```

```{r}
#Ridge Line Plot

# Step 2: subset: df without quality=NA

df_quality <- subset(df, !is.na(quality.group))

chemicals_group_by_qty <- list(
                                df_quality$alcohol,
                                df_quality$citric.acid,
                                df_quality$chlorides,
                                df_quality$density,
                                df_quality$fixed.acidity, 
                                df_quality$free.sulfur.dioxide, 
                                df_quality$pH,
                                df_quality$residual.sugar,
                                df_quality$sulphates,
                                df_quality$volatile.acidity,
                                df_quality$total.sulfur.dioxide
                             )

```

```{r}
#Ridge Line Plot

# Step 3: Function: Ridge Line Plot

create_ridgeline <- function(chemical, chemical_name, start, end, step){
  p <- ggplot(aes(x = chemical, y = quality.group, fill = quality.group), 
              data= df_quality) +
         geom_density_ridges(scale = 0.8, size = 0.25, rel_min_height = 0.03) +
         scale_x_continuous(limits = c(start, end),
                            breaks = seq(start, end, step))+
         theme_ridges() + 
         theme(legend.position = "none")+
         labs(
            x = chemical_name,
            y = "Quality"
         ) +
         my_theme
              
  return(p)
}
```


```{r}
#Violin Plot: Step 1

# Function to produce summary statistics (mean and +/- sd)

data_summary <- function(x) {
   m <- mean(x)
   ymin <- m-sd(x)
   ymax <- m+sd(x)
   return(c(y = m, ymin = ymin, ymax = ymax))
}

### Source: http://www.sthda.com/english/wiki/ggplot2-violin-plot-quick-start-guide-r-software-and-data-visualization
```

```{r}
#Violin Plot: Step 2

# Function: draw plot (chemicals vs quality.group)

create_violin <- function(chemical, chemical_name, start, end, step){
  p <-  ggplot(df_quality, 
               aes(x = quality.group, 
                   y = chemical, 
                   fill = quality.group)) + 
          geom_violin() + 
          scale_y_continuous(limits = c(start, end),
                                breaks = seq(start, end, step))+
          stat_summary(fun.data=data_summary)+ 
          coord_flip() + #flip x and y
          labs(
                  y = chemical_name,
                  x = "Quality"
          )+
          my_theme
            
  return(p)
}

# Because I flipped x and y,
# scale_y_continuous was used, and x- / y- labs were exchanged.

```


```{r results='hide'}
# Scatter Plot: 3 variables -- Two chemicals + Quality (group by alcohol)

# Step 1: group df by alcohol

# add a new variable for grouping
df$alcohol.group <-
  cut(df$alcohol, c(8.4,9.2,9.5,10,10.5,11,12,14.9))

# check the new variable
table(df$alcohol.group, useNA = 'ifany')

```

```{r}
# Step 2: build new subset without NA of alcohol

df_alcohol <- subset(df, !is.na(alcohol.group))
```

```{r}
# Step 3: Draw Plot

create_scatter_3var <- function(c1, c1_name, c1_unit, # c1 - Chemical at x-axis
                                start, end, step, 
                                c2, c2_name, c2_unit  # c2 - Chemical put into group
                                ) {
  p <- ggplot(aes(x = c1, y = quality, color = c2), data = df_alcohol) +
    
          geom_point(alpha = 0.5, size = 1, position = 'jitter') + 
    
          scale_color_brewer(type = 'div',
                       guide = guide_legend(title = paste(c2_name,'\n', c2_unit), 
                                            reverse = TRUE,
                                            override.aes = list(alpha = 1, size = 2)))+
    
          scale_x_continuous(
                             limits = c(start, end),
                             breaks = seq(start, end, step)
                             )+
          line_regression +
    
          labs(
                title = paste('Quality v.s.\n', c1_name, '&', c2_name),
                x = paste(c1_name, c1_unit),
                y = "Quality"
          ) +
    
          my_theme # theme()

  return(p)
  
}
```

```{r}
# Bivarite Plots: scatter + ridgeline + violine plots

draw_3_bivarite_plots <- function(c1, c2, c_name, start, end, step, unit){
  #c1 <- chemical
  #c2 <- chemicals_group_by_qty
  
  # step 1: draw plots
  p1 <- create_scatter(c1, c_name, start, end, step, unit)
  p2 <- create_ridgeline(c2, c_name, start, end, step)
  p3 <- create_violin(c2, c_name, start, end, step)
  
  # step 2: print plots
  print(p1)
  grid.arrange(p2, p3, nrow = 1)
}
```


```{r}
# The doughnut function permits to draw a donut plot
create_doughnut <-
function (x, labels = names(x), 
          edges = 200, 
          outer.radius = 0.8, 
          inner.radius=0.6, # inner.radius controls the width of the ring!
          clockwise = FALSE,
          init.angle = if (clockwise) 90 else 0, density = NULL, 
          angle = 45, col = NULL, border = FALSE, lty = NULL, 
          main = NULL, ...)
{
    if (!is.numeric(x) || any(is.na(x) | x < 0))
        stop("'x' values must be positive.")
    if (is.null(labels))
        labels <- as.character(seq_along(x))
    else labels <- as.graphicsAnnot(labels)
    x <- c(0, cumsum(x)/sum(x))
    dx <- diff(x)
    nx <- length(dx)
    plot.new()
    pin <- par("pin")
    xlim <- ylim <- c(-1, 1)
    if (pin[1L] > pin[2L])
        xlim <- (pin[1L]/pin[2L]) * xlim
    else ylim <- (pin[2L]/pin[1L]) * ylim
    plot.window(xlim, ylim, "", asp = 1)
    if (is.null(col))
        col <- if (is.null(density))
          palette()
        else par("fg")
    col <- rep(col, length.out = nx)
    border <- rep(border, length.out = nx)
    lty <- rep(lty, length.out = nx)
    angle <- rep(angle, length.out = nx)
    density <- rep(density, length.out = nx)
    twopi <- if (clockwise)
        -2 * pi
    else 2 * pi
    t2xy <- function(t, radius) {
        t2p <- twopi * t + init.angle * pi/180
        list(x = radius * cos(t2p), 
             y = radius * sin(t2p))
    }
    for (i in 1L:nx) {
        n <- max(2, floor(edges * dx[i]))
        P <- t2xy(seq.int(x[i], x[i + 1], length.out = n),
                  outer.radius)
        polygon(c(P$x, 0), c(P$y, 0), density = density[i], 
                angle = angle[i], border = border[i], 
                col = col[i], lty = lty[i])
        Pout <- t2xy(mean(x[i + 0:1]), outer.radius)
        
        # labels
        lab <- as.character(labels[i])
        
        if (!is.na(lab) && nzchar(lab)) {
            lines(c(1, 1.05) * Pout$x, c(1, 1.05) * Pout$y)
            text(1.1 * Pout$x, 1.1 * Pout$y, labels[i], 
                 xpd = TRUE, adj = ifelse(Pout$x < 0, 1, 0), 
                 ...)
        }
        ## Add white disc          
        Pin <- t2xy(seq.int(0, 1, length.out = n*nx),
                  inner.radius)
        polygon(Pin$x, Pin$y, density = density[i], 
                angle = angle[i], border = border[i], 
                col = "white", lty = lty[i])
    }
 
    title(main = main, ...)
    invisible(NULL)
}


# source: https://www.r-graph-gallery.com/130-ring-or-donut-chart/
 
# Let's use the function, it works like PiePlot !
# inner.radius controls the width of the ring!

# test
# create_doughnut(qlt_count, inner.radius=0.5,labels = qlt_category,col=qlt_color)

```


```{r}
text_bivar <- list(
#alcohol  1
'Alcohol and Quality

From scatter plot, it can be seen that 
when alcohol increases, the quality score rises.

From the ridgeline and violine plot, it could be found that 
when the quality score increases, 
the alcohol distribution moves to the right (alcohol increases), 
but the shapes are different to each other. 

Thus, I guess even though they are positively correlated, 
the correlation might not be strong.

The spearman coefficient (0.48) shows that they are mild correlated, 
same as what we got from the plots.
',

#citric.acid 2
'Citric Acid and Quality

From scatter plot, it can be seen that 
the quality score does not have any changes,
when the citric acid increases.

In the ridgeline and violine plot,
the x of the distributions does not move,
with the change of the quality groups.
The shape of the 2 groups in the middle are almost same.

Thus, I guess Citric Acid and Quality are not correlated.

The spearman coefficient is 0.21 (not correlated),
indicating a same result as what we got from the plots.
',

#chlorides 3
'Chlorides and Quality

From scatter plot, it can be seen that 
the quality score does not have any changes,
when the cchlorides increase.

In the ridgeline and violine plot,
the x of the distributions just move to the left a bit,
with the change of the quality groups.
The shape of the 4 groups in the plots are very similar, 
especially the 2 groups in the middle, which represents most of the wines.

Thus, I guess chlorides and Quality are not correlated.

The spearman coefficient is -0.19 (not correlated)indicating a same result as what we got from the plots.
',

#density 4
'Density and Quality 

From scatter plot, it can be seen that 
the quality score does not have any changes,
when the density increase.

In the ridgeline and violine plot,
the x of the distributions does not move,
with the change of the quality groups.
The shape of the 4 groups in the plots are very similar, 
especially the 2 groups in the middle, which represents most of the wines.

Thus, I guess density and Quality are not correlated.

The spearman coefficient is -0.18 (not correlated),
indicating a same result as what we got from the plots.
',

#fixed.acidity 5
'Fixed Acidity and Quality 

From scatter plot, it can be seen that 
the quality score does not have any changes,
when the fixed acidity increases.

In the ridgeline and violine plot,
the x of the distributions does not move,
with the change of the quality groups.
The shape of the 2 groups in the middle are almost same.

Thus, I guess fixed acidity  and Quality are not correlated.

The spearman coefficient is 0.11 (not correlated),
indicating a same result as what we got from the plots.
', 

#free.sulfur.dioxide 6 
'Free Sulfur Dioxide and Quality 

From scatter plot, it can be seen that 
the quality score does not have any changes,
when Free Sulfur Dioxide increase.

In the ridgeline and violine plot,
the x of the distributions does not move,
with the change of the quality groups.
The shape of the 4 groups in the plots are very similar.

Thus, I guess Free Sulfur Dioxide and Quality are not correlated.

The spearman coefficient is -0.057 (not correlated),
indicating a same result as what we got from the plots.
', 

#pH 7
'PH values and Quality

From scatter plot, it can be seen that 
the quality score does not have any changes,
when pH increase.

In the ridgeline and violine plot,
the x of the distributions does not move,
with the change of the quality groups.
The shape of the 4 groups in the plots are very similar.

Thus, I guess pH and Quality are not correlated.

The spearman coefficient is -0.044 (not correlated),
indicating a same result as what we got from the plots.
', 

#residual.sugar 8 
'Residual Sugar and Quality 

From scatter plot, it can be seen that 
the quality score does not have any changes,
when Residual Sugar increase.

In the ridgeline and violine plot,
the x of the distributions does not move,
with the change of the quality groups.
The shape of the 3 groups at the top are very similar.

Thus, I guess Residual Sugar and Quality are not correlated.

The spearman coefficient is -0.031 (not correlated),
indicating a same result as what we got from the plots.
', 
         
         
#sulphates 9
'Sulphates and Quality

From scatter plot, it can be seen that 
when Sulphates < 0.9, the quality score rises as Sulphates increase. 
When Sulphates > 0.9, the quality score drops as Sulphates increase. 

From the ridgeline and violine plot, it could be found that 
when the quality score increases, 
the Sulphates distribution moves to the right (Sulphates increase).
Besides, the shapes are similar to each other. 

Thus, I guess even though they are positively correlated, 
the correlation might not be strong.

The spearman coefficient (0.39) shows that they are mild correlated.

It might be a better way to use normalised Sulphates data, or discuss the relationship by grouping the Sulphates.
',

#volatile.acidity 10
'Volatile Acidity and Quality

From scatter plot, it can be seen that 
the quality score decreases, when Volatile Acidity increase,
but the trend is not obvious.

In the ridgeline and violine plot,
the x of the distributions moves to the left,
with the change of the quality groups.
The shape of the 4 groups are similar.

Thus, I guess Residual Sugar and Quality are slightly negatively correlated.

The spearman coefficient is -0.38 (weak negative correlation),
indicating a same result as what we got from the plots.
', 

#total.sulfur.dioxide 11
'Free Sulfur Dioxide and Quality

From scatter plot, it can be seen that 
the quality score does not change, when Volatile Acidity increase.

In the ridgeline and violine plot,
the x of the distributions moves to the left,
with the change of the quality groups.
The shape of the 4 groups are similar.

Thus, I guess Free Sulfur Dioxide and Quality are not correlated.

The spearman coefficient is -0.2 (not correlated),
indicating a same result as what we got from the plots.
'
)
```

```{r}
text_univar <- list(
  #alcohol  1
  
'The percentage of alcohol is between 8.40 and 14.90.\n
It is rightly skewed, and most wines have 9% -12% alcohol,
with the median of 10.20% and mean of 10.42%.\n
I tried to normalise it by log10 or sqrt, but the distribution of them did not seem better.',
  
  #citric.acid 2
  
'Citric Acid ranges from 0 to 1 g / dm^3.\n
There are 3 peaks in the distribution of Citric Acid with a long tail to the right.\n
The 1st peak is between 0.0-0.1, with the 2nd in 0.2-0.3 and 3rd in 0.4-0.5.\n
There is a sharp decrease betwenn 0.5 and 0.6. \n
Most of wines have less than 0.5 g / dm^3 Citric Acid.\n
Again, the log10 and sqrt method could not normalise it.',
  
#chlorides 3
'The distribution of Chlorides is almost normal.\n
Most wines have 0.04 - 0.12 g / dm^3 Chlorides, with the mean of 0.087.\n
Though it ranges from 0.012 to 0.611,
there are very few wines with the Chlorides of 0.012 - 0.04 or 0.12 - 0.611.',

#density 4
'Density is almost normaly distributed.\n
Most wines have the density of 0.9956 - 0.9978 g / dm^3, 
with the mean of 0.9967.',

#fixed.acidity 5
'Fixed Acidity is a bit rightly skewed.\n
Though it ranges from 4.60 to 15.90 g / dm^3,
most of wines have the fixed acidity of 5 -13, 
with the median of 7.90. \n
The mode is between 7 -8.\n
Compared with sqrt, Log10 seems to be a better method to nomalise it.', 

#free.sulfur.dioxide 6 
'Free Sulfur Dioxide is right skewed with a long tail to the right.\n
The peak of the distribution is between 6 - 11 mg/dm^3. \n
Though it ranges from 1 - 72, 
most wines have the Free Sulfur Dioxide less than 41.\n
Sqrt could make it like a normal distribution than log10.', 

#pH 7
'PH values look like normally distributed. \n
It mainly ranges from 3.0 to 3.6, with the mean of 3.311.', 

#residual.sugar 8 
'Residual Sugar has a long tail to the right.\n 
Though it ranges from 0.9 to 15.5 g/dm^3,
the sugar of most wines are less than 3.0.\n
If we cut it from 3.0, the distribution less than 3.0 is almost normal.\n
Log10 method is more suitable to normalise the Sulphates', 
              
#sulphates 9
'Sulphates is right skewed.\n
Though it ranges from 0.33 to 2.00 g/dm^3,
the Sulphates of most wines are less than 1.0.\n
From the histogram, it could be seen that
there is another small peak between these two (near 0.5).\n
Log10 method is more suitable to normalise the Sulphates',

#volatile.acidity 10
'Volatile is bimodal distributed, 
with the 1st peak around 0.4 g/dm^3 and the 2nd near 0.6.\n
Most wines have the acidity between 0.3 to 0.7.', 

#total.sulfur.dioxide 11
'Free Sulfur Dioxide is right skewed.\n
Most wines have the Free Sulfur Dioxide less than 156 mg / dm^3, 
with a few outliers reaching 289.\n
Log10 could well nomalise it.'
)
```



## Univariate

The main purpose of this report is to discover the chemicals that could predict the quality of wines. Besides, from the name of the 12 variables, we may find that some of them might be correlated with others (e.g. Acid and PH).

Thus, the analysis would include the calculation of correlation and linear regression, both of which may requires a normal distribution of variables.

In the univariate plots and analysis, it is focused on exploring the distribution of variables.

### Histogram of All the Variables
There are 11 continuous variables and one ordinal variable. Most of them are skewed to right.
```{r}
# Show the histogram of all variables

# Draw histograms
df_s %>%  # Drop 1st column 
  as.tibble() %>%
  gather(variable, value) %>%
  ggplot( aes(x=value) ) +
    geom_histogram(alpha = my_alpha, fill= my_color) +
    facet_wrap(~variable, scale="free")
```

### Quality (Rating 0 - 10)
Quality is the only ordinal variable in the data set. Though the range of it is 0-10, there is no wine ranked as 0-2 or 9-10. Most of the wines were scored as  5-6.

```{r}
summary(df$quality)
```

```{r results = 'hide'}
table(df$quality)
```

```{r}
# Ordinal Variable: use 'geom_bar()' instead of 'geom_histogram()'

ggplot(df, aes(x=factor(quality))) + 
  geom_bar(alpha = my_alpha, fill= my_color) +
  labs(
    x = 'Quality Ratings',
    y = 'Count'
  )
  
```


```{r}
# Donut Plot args

qlt_category <- c(3, 4, 5, 6, 7, 8) #6 categories

qlt_count <- c(10, 53, 681, 638, 199, 18)

qlt_color <- c(blue[[5]], red[[2]], blue[[1]], blue[[2]], blue[[3]], red[[1]]) #6 colors

```


```{r}
# Draw donut plot

create_doughnut(qlt_count, 
              inner.radius=0.5, # inner.radius controls the width of the ring!
              labels = qlt_category,
              col=qlt_color )
```

### Chemicals
```{r}
# Draw Univariate Plots (group of 4)
for (i in 1:11) {

  # print the 4 plots
  p <- create_4_plots_table(
                            chemicals[[i]],
                            chemical_names[[i]],
                            chemical_args[[i]][[1]],
                            chemical_args[[i]][[2]],
                            chemical_args[[i]][[3]],
                            chemical_units[[i]]
                          )
  
  #print(p) -- ggplot requires 'print(p)' as it is in this loop function, but no need 'print(p)' with arrange.grib. Otherwise the output includes the gtable.
  
  # summary of the variable
  s <- summary(chemicals[[i]])
  
  # print markdown title: cat()
  cat(paste(chemical_names[[i]], '\n\n\n'))
  print(s)
  cat('\n\n')
  cat(text_univar[[i]])
}
```




## Bivariate - Quality vs Chemicals

### Plots: Quality vs Chemicals
```{r}
# 3 kinds of plot to explore 
# the relationship between a chemical and the quality

for (i in 1:11){
  
  # step 1: get arguments
  c1 <- chemicals[[i]]
  c2 <- chemicals_group_by_qty[[i]]
  
  c_name <- chemical_names[[i]]
  
  unit <- chemical_units[[i]]
  
  start <- chemical_args[[i]][[1]]
  end <- chemical_args[[i]][[2]]
  step <- chemical_args[[i]][[3]]
  
  # step 2: draw plots
  draw_3_bivarite_plots(c1, c2, c_name, start, end, step, unit)
  
  # step 3: add analysis
  cat(text_bivar[[i]])
}
```

From the plots above, we can find that there are 3 chemicals correlated with quality:

+ Volatile Acidity: negative
+ Sulphates: positive
+ Alcohol: positive

The following section summarises all of the Spearman Correlation Coefficients mentioned above.

### Spearman Correlation Coefficient: Quality vs Chemicals
```{r}
# Spearman Correlation: Quality vs Chemicals

# Build a new data frame to store r
cor_quality <- data.frame(matrix(nrow = 11, ncol = 2))

# Name 2 columns of table
colnames(cor_quality) <- c("chemicals", "r")

# Put r into the new data frame
for (i in 1:11) {
 cor <- cor(chemicals[[i]], df$quality, method = "spearman")
 cor_quality [i, 1] <- chemical_names[i]
 cor_quality [i, 2] <- cor
}
```

```{r}
# Show r in table

# Ordered by absolute value of r
cor_table <- arrange(cor_quality, desc(abs(r)))

# Print in table format
cor_table %>% kable() %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed", "responsive"), 
    full_width = F
  )

```

```{r}
# Plot of r

# Reorder
cor_quality %>%
  arrange(abs(r)) %>%
  #mutate(new_var = existing_var)
  mutate(c = factor(chemicals, chemicals)) %>%
  ggplot(aes(x = c, y = r)) +
    geom_segment( aes(x = c, xend=c, y=0, yend=r), 
                  color="skyblue", 
                  size=1) +
    geom_point( color="blue", size=4, alpha=0.6) +
    theme_light() +
    coord_flip() +
    theme(
      panel.grid.major.y = element_blank(),
      panel.border = element_blank(),
      axis.ticks.y = element_blank()
    ) +
  xlab("") +
  ylab("Spearman Correlation Coefficience with Quality")
```

+ Alcohol, Volatile Acidity and Sulphates are correlated with quality.

+ Other chemicals are not correlated with quality.



## Bivariate - Chemicals
After analysing the relationship between the quality score and chemicals,
the relationship between 11 chemicals will be explored in this section.

```{r results = 'hide'}

# Compute 'Pearson' correlation between chemicals

df_no_quality <- df[, (2:12)]
r <- cor(df_no_quality)
head(round(r,2))
```

### Correlation Matrix
```{r}
# Build a Pannel of 100 colors with Rcolor Brewer
r_colors <- brewer.pal(5, "RdBu")
r_colors=colorRampPalette(r_colors)(100)
 
# Order the correlation matrix
ord <- order(r[1, ])
r_ord = r[ord, ord]
plotcorr(r_ord , col=r_colors[r_ord*50+50] , mar=c(1,1,1,1)  )
```

From the correlation matrix, it could be seen that there some chemicals are highly correlated with others:

+ pH v.s Fixed Acidity
+ pH v.s. Citric Aciity
+ Citric Acid v.s. Fixed Acidity
+ Citric Acidity v.s. Volatile Acidity
+ Density v.s. Fixed Acidity
+ Free Sulfur Dioxide v.s. Total Sulfur Dioxide

Besides, the three chemicals that are correlated with quality
- Alcohol, Volatile Acidity and Sulphates, are not correlate with each other.

The Pearson correlation coefficients between chemicals is shown in the following table.

### Pearson Correlation Coefficient between Chemicals
```{r results='hide'}
#To compute the matrix of p-value, a custom R function is used :

# mat : is a matrix of data
# ... : further arguments to pass to the native R cor.test function
cor.mtest <- function(mat, ...) {
    mat <- as.matrix(mat)
    n <- ncol(mat)
    p.mat<- matrix(NA, n, n)
    diag(p.mat) <- 0
    for (i in 1:(n - 1)) {
        for (j in (i + 1):n) {
            tmp <- cor.test(mat[, i], mat[, j], ...)
            p.mat[i, j] <- p.mat[j, i] <- tmp$p.value
        }
    }
  colnames(p.mat) <- rownames(p.mat) <- colnames(mat)
  p.mat
}
# matrix of the p-value of the correlation
p.mat <- cor.mtest(df_no_quality)
head(p.mat[, 1:5])

### Source: http://www.sthda.com/english/wiki/visualize-correlation-matrix-using-correlogram
```

```{r}
col <- colorRampPalette(c("#BB4444", "#EE9988", "#FFFFFF", "#77AADD", "#4477AA"))

corrplot(r, method="color", col=col(1000),  
         type="upper", order="hclust", 
         addCoef.col = "black", # Add coefficient of correlation
         tl.col="black", tl.srt=20, # Text label color and rotation
         p.mat = p.mat, sig.level = 0.05, insig = "blank", # Combine with significance
         diag=FALSE,  # hide correlation coefficient on the principal diagonal
         number.cex=0.75, # change font size of r
         tl.cex = 0.8 # change font size of labels
         )

### source: http://www.sthda.com/english/wiki/visualize-correlation-matrix-using-correlogram
```

Pearson Correlation Coefficient of Highly Correlated Chemicals:

+ pH v.s Fixed Acidity: -0.68
+ pH v.s. Citric Aciity: -0.54
+ Citric Acid v.s. Fixed Acidity: 0.67
+ Citric Acid v.s. Volatile Acidity: -0.55
+ Density v.s. Fixed Acidity: 0.67
+ Free Sulfur Dioxide v.s. Total Sulfur Dioxide:  0.67

The relationship between these chemicals are shown in the following plots.

### Moderately / Highly Correlated Chemicals
```{r}
# plot the chemicals with r>0.5
scatterplotMatrix(~citric.acid +
                    density +
                    fixed.acidity + 
                    free.sulfur.dioxide + 
                    pH +
                    total.sulfur.dioxide, 
                  data = df ,
                  reg.line = FALSE,
                  col = my_color , 
                  pch = 20,
                  cex =1.5,
                  main = "Scatter Plot of Correlated Chemicals" )

```

## Multivariate - Chemicals

### PH, Fixed Acidity & Density
From the correlation matrix, it is shown that PH, Fixed Acidity & Density are highly correlated with each other. Thus, the following plot is draw to present the relationship between them.

```{r results='hide'}
## Step 1: group df by fixed.acidity

# add a new variable for grouping
df$fixed.acidity.group <-
  cut(df$fixed.acidity, c(4.60, 7.10, 7.90, 9.20, 15.90))

# check the new variable
table(df$fixed.acidity.group, useNA = 'ifany')
```

```{r}
# Step 2: build new subset without NA of fixed.acidity

df_fixed.acidity <- subset(df, !is.na(fixed.acidity.group))
```

```{r}
# Step 3: draw plot - PH, Fixed Acidity & Density
p_3 <- ggplot(aes(x = citric.acid, y = pH, color = fixed.acidity.group),
              data = df_fixed.acidity )+
        geom_point(alpha = 0.5, size = 1, position = 'jitter')+
        scale_color_brewer(type = 'div',
                           guide = guide_legend(title = "Fixed Acidity\n (g / dm^3)", 
                                                reverse = TRUE,
                                                override.aes = list(alpha = 1, size = 2)))+
        scale_x_continuous()+
        ggtitle('PH Values v.s.\n Citric Acid & Fixed Acidity')+ 
        xlab("Citric Acid (g / cm^3)") + 
        ylab("PH Value")+
        my_theme +
        line_regression

print(p_3)
```

From the plot, we can easily see the trends:

+ Citric Acid and pH values are negatively correlated.

+ Fixed acidity and pH values are negatively correlated.

+ Fixed acidity and Citric Acid are positively correlated.




## Multivariate - Quality v.s. Chemicals

3 chemicals are correlated with quality:

+ Volatile Acidity: negative
+ Sulphates: positive
+ Alcohol: positive

From the correlation matrix above, it can be found that these 3 chemicals are not correlated with each other. Thus, they independently affect quality scores.

### Quality, Alcohol and Volatile Acidity
```{r}
# Draw Scatter Plot: volatile.acidity & alcohol & quality
p_1 <- create_scatter_3var(df_alcohol$volatile.acidity, 'Volatile Acidity', 
                    chemical_units[[10]],
                    chemical_args[[10]][[1]],
                    chemical_args[[10]][[2]],
                    chemical_args[[10]][[3]],
                    df_alcohol$alcohol.group, 'Alcohol',
                    chemical_units[[1]])
print(p_1) # p_1 will be used in "Final Plots" Section
```

+ Alcohol and quality are positively correlated.

+ Volatile acidity and quality are negatively correlated.

### Quality, Alcohol and Sulphates
```{r}
# Draw Scatter Plot: sulphates & alcohol & quality
p_2 <- create_scatter_3var(df_alcohol$sulphates, 'Sulphates', 
                    chemical_units[[9]],
                    chemical_args[[9]][[1]],
                    chemical_args[[9]][[2]],
                    chemical_args[[9]][[3]],
                    df_alcohol$alcohol.group, 'Alcohol',
                    chemical_units[[1]]
                    )
print(p_2) # p_2 will be used in "Final Plots" Section
```

+ Alcohol and quality are positively correlated.

+ Sulphates and quality are positively correlated.



# Final Plots and Summary

## Quality, Alcohol and Volatile Acidity
```{r}
# Draw Scatter Plot: volatile.acidity & alcohol & quality
print(p_1)
```

+ Alcohol and quality are positively correlated.

+ Volatile acidity and quality are negatively correlated.


## Quality, Alcohol and Sulphates
```{r}
# Draw Scatter Plot: sulphates & alcohol & quality
print(p_2)
```

+ Alcohol and quality are positively correlated.

+ Sulphates and quality are positively correlated.


## PH Values, Density and Fixed Acidity
```{r}
# Scatter plot: PH, Fixed Acidity & Density
print(p_3)
```

+ Density and pH values are negatively correlated.

+ Fixed acidity and pH values are negatively correlated.

+ Fixed acidity and density values are positively correlated.


# Reflection

## Conclusion
3 out of 11 chemicals are moderately correlated with quality:

+ Volatile Acidity (negatively),
+ Sulphates (positively) and
+ Alcohol (positively),

while others are not correlated.

Besides, these 3 chemicals are not correlated with each other. Thus, they can be used to predict the quality score.

## Limitation

There are some chemicals highly correlated with each other:

+ pH and Fixed Acidity
+ pH and. Citric Aciity
+ Citric Acid and Fixed Acidity
+ Citric Acidity and Volatile Acidity
+ Density and Fixed Acidity
+ Free Sulfur Dioxide and Total Sulfur Dioxide

To design a better experiment, it would be helpful to drop some of the highly correlated variables.


Besides, the quality scores are mainly between 5-6, and some are 7. It can be noticed that none of the wines were scored 0-2 and 9-10. Usually a 5-point scale would be better for scoring. 

It might also be helpful to ask experts to score the acidity and sweetness after tasting (e.g 1-Not sweet at all ~ 5-Too sweet). Then we can compare the degree of real sweetness /sourness and the subjective feeling of it.

## Success and Improvement

I feel it is much more clear to separate the data and the plot function. The code structure is better after gathering data and put it into array-like storage.

I would try more kinds of plots and to add more details for the plots the next time (e.g. I tried to add a layer to show correlation coefficiency on the scatter plot, but I failed.)


[^1]: Information of the Original Data:
 P. Cortez, A. Cerdeira, F. Almeida, T. Matos and J. Reis. 
 Modeling wine preferences by data mining from physicochemical properties.
 In Decision Support Systems, Elsevier, 47(4):547-553. ISSN: 0167-9236.
