# Red Wine Quality Analysis
### Which chemical properties influence the quality of red wines?

By Linjing Yang

# Introduction
The tidy data set contains 1,599 red wines with 11 variables on the chemical properties of the wine. At least 3 wine experts rated the quality of each wine, providing a rating between 0 (very bad) and 10 (very excellent).

+ 1599 records
+ 11 chemical properties
+ quality rating: 0 - 10

## Variables

###Input variables (based on physicochemical tests):
1. fixed acidity (tartaric acid - g / dm^3)
2. volatile acidity (acetic acid - g / dm^3)
3. citric acid (g / dm^3)
4. residual sugar (g / dm^3)
5. chlorides (sodium chloride - g / dm^3
6. free sulfur dioxide (mg / dm^3)
7. total sulfur dioxide (mg / dm^3)
8. density (g / cm^3)
9. pH
10. sulphates (potassium sulphate - g / dm3)
11. alcohol (% by volume)
   
###Output variable (based on sensory data): 
12. quality (score between 0 and 10)



#### Information of the Original Data:
+  P. Cortez, A. Cerdeira, F. Almeida, T. Matos and J. Reis. 
+  Modeling wine preferences by data mining from physicochemical properties.
+  In Decision Support Systems, Elsevier, 47(4):547-553. ISSN: 0167-9236.


# Exploration of Data
```{r echo=FALSE}
# {r echo=FALSE} Do not show this code block.

# Avoid all codes and warnings to be shown in the HTML document
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)
```

```{r}
### Load libraries
library(GGally)
library(ggplot2)
library(gridExtra)
library(RColorBrewer) # Scatter Plot: color
library(dplyr) # %>%
library(tidyr)
library(memisc) # Linear Regression: mtable
library(corrplot) # correlation matrix
library(ggridges) # Ridge Line Plot
library(kableExtra) # Show the head of the orginal data
library(tibble) # Check distribution of the whole group of chemicals
library(car) # Scatterplot Matrix
library(corrplot) # Correlation Matrix (with cor. number)

library(ellipse)
library(RColorBrewer)
```

```{r results = 'hide'}
#{r results = 'hide'} Do not show the result of this code block.

### Load the data
getwd()
list.files()
df <- read.csv('wineQualityReds.csv')
head(df)
```

```{r results = 'hide'}
# Details of data
str(df)
```

```{r}
# show the head of the data
df[, (2:13)] %>% head(6) %>% kable() %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed", "responsive"), 
    full_width = F
  )
```

```{r}
# All chemicals in df
chemicals <-  list(
                df$alcohol,
                df$citric.acid,
                df$chlorides,
                df$density,
                df$fixed.acidity, 
                df$free.sulfur.dioxide, 
                df$pH,
                df$residual.sugar,
                df$sulphates,
                df$volatile.acidity,
                df$total.sulfur.dioxide
                  )

# List:
#   Index of lists starts from [1] rather than [0]
#   Access the element of a list: list[[index]]

```

```{r}
# Names of all chemicals
chemical_names <- c(
    'Alcohol',
    'Citric Acid',
    'Chlorides',
    'Density',
    'Fixed Acidity', 
    'Free Sulfur Dioxide', 
    'pH Value',
    'Residual Sugar',
    'Sulphates',
    'Volatile Acidity',
    'Total Sulfur Dioxide'
)
```

```{r}
# Plot arguments of chemicals

# c(start, end, step)

chemical_units <- list(
                      #alcohol  1
                      '(% by volume)',
                      
                      #citric.acid 2
                      '(g / dm^3)',
                      
                      #chlorides 3
                      '(g / dm^3)',
                      
                      #density 4
                      '(g / cm^3)',
                      
                      #fixed.acidity 5
                      '(g / dm^3)', 
                      
                      #free.sulfur.dioxide 6 
                      '(mg / dm^3)', 
                      
                      #pH 7
                      '', 
                      
                      #residual.sugar 8 
                      '(g / dm^3)', 
                      
                      #sulphates 9
                      '(g / dm3)',
                      
                      #volatile.acidity 10
                      '(g / dm^3)', 
                      
                      #total.sulfur.dioxide 11
                      '(mg / dm^3)' 
                    )
```

```{r}
# Plot arguments of chemicals

# c(start, end, step)

chemical_args <- list(
                      #alcohol  1
                      c(8, 15, 1),
                      
                      #citric.acid 2
                      c(0, 1, 0.1),
                      
                      #chlorides 3
                      c(0, 0.2, 0.02),
                      
                      #density 4
                      c(0.99, 1.0037, 0.01),
                      
                      #fixed.acidity 5
                      c(4, 16, 1), 
                      
                      #free.sulfur.dioxide 6 
                      c(1, 72, 10), 
                      
                      #pH 7
                      c(2.7, 4.1, 0.5), 
                      
                      #residual.sugar 8 
                      c(1, 5, 0.5), 
                      
                      #sulphates 9
                      c(0.3, 1.3, 0.2),
                      
                      #volatile.acidity 10
                      c(0.1, 1.2, 0.2), 
                      
                      #total.sulfur.dioxide 11
                      c(6, 160, 50) 
                    )
```

```{r}
# Function: Density plot

create_density <- 
  function(chemical, chemical_name, start, end, step, chemical_unit){
    p <- ggplot(aes(x=chemical), 
                data =df) +
         geom_density(alpha = my_alpha, fill= my_color, color = my_color) +
         scale_x_continuous(limits = c(start, end),
                            breaks = seq(start, end, step))+
         labs(
           x = paste(chemical_name, chemical_unit),
           y = 'Density')
    return(p)
  }
```

```{r}
# Function: create 3 histograms: orginal, log10, sqrt
# And put density plot together

create_4_plots_table <- 
  function(chemical, chemical_name, start, end, step, chemical_unit){
    p1 <- ggplot(aes(x=chemical), data= df) +
            geom_histogram(alpha = my_alpha, fill= my_color) +
            xlab(paste(chemical_name, chemical_unit, sep = ' '))
    
    p2 <- p1 + scale_x_log10() + 
               xlab(paste(chemical_name, "(log10)", sep = ' '))
    
    p3 <- p1 + scale_x_sqrt() + 
               xlab(paste(chemical_name, "(sqrt)", sep = ' '))
  
    p0 <- create_density(chemical, chemical_name, 
                         start, end, step, chemical_unit)
    
    p <- grid.arrange(p0, p1, p2, p3, 
                      ncol=2, 
                      top = chemical_name # Name of the four plots
                      ) 
    return(p)
  }
```

```{r}
# Scatter Plot with Lines

# Step 1: set arguments

my_color <- '#008AD8'   # Main Color of this report: blue

my_alpha <- 0.7

line_mean <- geom_line(stat='summary', fun.y=mean, 
                       linetype=1, color='red')

line_median <- geom_line(stat='summary', fun.y=quantile,
                         fun.args=list(probs=.5),
                         linetype=1, color=my_color)

my_theme <- theme(
                plot.title = element_text(size = 20),  
                axis.title.x = element_text(size = 14),  
                axis.title.y = element_text(size = 14),
                legend.title = element_text(size = 14),
                legend.text = element_text(size = 12)
             )
```

```{r}
# Scatter Plot with Lines

# Step 2: Draw Plot
create_scatter <- 
  function(chemical, chemical_name, start, end, step, chemical_unit){
    p <- ggplot(aes(x=chemical, y=quality),data=df)+
          geom_point(alpha = 0.1, size = 1, position = 'jitter') +
          scale_x_continuous(limits = c(start, end),
                            breaks = seq(start, end, step)) +
          line_mean +
          line_median +
          labs(
              x = paste(chemical_name, chemical_unit),
              y = "Quality"
           )
    
    return(p)
  }
```


```{r results = 'hide'}
#Ridge Line Plot

# Step 1: group df by quality

df$quality.group <-
  cut(df$quality,
      c(3,4,5,6,8))

table(df$quality.group,
      useNA = 'ifany')
```

```{r}
#Ridge Line Plot

# Step 2: subset: df without quality=NA

df_quality <- subset(df, !is.na(quality.group))

chemicals_group_by_qty <- list(
                                df_quality$alcohol,
                                df_quality$citric.acid,
                                df_quality$chlorides,
                                df_quality$density,
                                df_quality$fixed.acidity, 
                                df_quality$free.sulfur.dioxide, 
                                df_quality$pH,
                                df_quality$residual.sugar,
                                df_quality$sulphates,
                                df_quality$volatile.acidity,
                                df_quality$total.sulfur.dioxide
                             )

```

```{r}
#Ridge Line Plot

# Step 3: Function: Ridge Line Plot

create_ridgeline <- function(chemical, chemical_name, start, end, step){
  p <- ggplot(aes(x = chemical, y = quality.group, fill = quality.group), 
              data= df_quality) +
         geom_density_ridges(scale = 0.8, size = 0.25, rel_min_height = 0.03) +
         scale_x_continuous(limits = c(start, end),
                            breaks = seq(start, end, step))+
         theme_ridges() + 
         theme(legend.position = "none")+
         labs(
            x = chemical_name,
            y = "Quality"
         ) +
         my_theme
              
  return(p)
}
```


```{r}
#Violin Plot: Step 1

# Function to produce summary statistics (mean and +/- sd)

data_summary <- function(x) {
   m <- mean(x)
   ymin <- m-sd(x)
   ymax <- m+sd(x)
   return(c(y = m, ymin = ymin, ymax = ymax))
}

### Source: http://www.sthda.com/english/wiki/ggplot2-violin-plot-quick-start-guide-r-software-and-data-visualization
```

```{r}
#Violin Plot: Step 2

# Function: draw plot (chemicals vs quality.group)

create_violin <- function(chemical, chemical_name, start, end, step){
  p <-  ggplot(df_quality, 
               aes(x = quality.group, 
                   y = chemical, 
                   fill = quality.group)) + 
          geom_violin() + 
          scale_y_continuous(limits = c(start, end),
                                breaks = seq(start, end, step))+
          stat_summary(fun.data=data_summary)+ 
          coord_flip() + #flip x and y
          labs(
                  y = chemical_name,
                  x = "Quality"
          )+
          my_theme
            
  return(p)
}

# Because I flipped x and y,
# scale_y_continuous was used, and x- / y- labs were exchanged.

```

```{r}
# test
#create_violin(df_quality$alcohol, 'Alcohol', 8, 15, 1)
```

```{r results='hide'}
# Scatter Plot: 3 variables -- Two chemicals + Quality (group by alcohol)

# Step 1: group df by alcohol

# add a new variable for grouping
df$alcohol.group <-
  cut(df$alcohol, c(8.4,9.2,9.5,10,10.5,11,12,14.9))

# check the new variable
table(df$alcohol.group, useNA = 'ifany')

```

```{r}
# Step 2: build new subset without NA of alcohol

df_alcohol <- subset(df, !is.na(alcohol.group))
```

```{r}
# Step 3: Draw Plot

create_scatter_3var <- function(c1, c1_name, c1_unit, # c1 - Chemical at x-axis
                                start, end, step, 
                                c2, c2_name, c2_unit) { # c2 - Chemical put into group
  p <- ggplot(aes(x = c1, y = quality, color = c2), data = df_alcohol) +
    
          geom_point(alpha = 0.5, size = 1, position = 'jitter') + 
    
          scale_color_brewer(type = 'div',
                       guide = guide_legend(title = paste(c2_name, c2_unit), 
                                            reverse = TRUE,
                                            override.aes = list(alpha = 1, size = 2)))+
    
          scale_x_continuous(
                             limits = c(start, end),
                             breaks = seq(start, end, step)
                             )+
          line_mean +
          line_median +
    
          labs(
                title = paste('Quality v.s.\n', c1_name, '&', c2_name),
                x = paste(c1_name, c1_unit),
                y = "Quality"
          ) +
    
          my_theme # theme()

  return(p)
  
}
```


## Univariate Plots

### Histogram of All the Variables
```{r}
# Show the histogram of all variables

# Draw histograms
df[, (2:13)] %>%  # Drop 1st column 
  as.tibble() %>%
  gather(variable, value) %>%
  ggplot( aes(x=value) ) +
    geom_histogram(alpha = my_alpha, fill= my_color) +
    facet_wrap(~variable, scale="free")
```


### Quality (Rating 0 - 10)

```{r}
summary(df$quality)
```

```{r results = 'hide'}
table(df$quality)
```

```{r}
ggplot(aes(x=quality, y = ..count..), data =df) +
  geom_histogram(alpha = my_alpha, fill= my_color)+
  scale_x_continuous(limits = c(0,10),
                     breaks = seq(0, 10,1))
```



### Chemicals
```{r}
# Draw Univariate Plots (group of 4)
for (i in 1:11) {

  # print the 4 plots
  p <- create_4_plots_table(
                            chemicals[[i]],
                            chemical_names[[i]],
                            chemical_args[[i]][[1]],
                            chemical_args[[i]][[2]],
                            chemical_args[[i]][[3]],
                            chemical_units[[i]]
                          )
  
  #print(p) -- ggplot requires 'print(p)' as it is in this loop function, but no need 'print(p)' with arrange.grib. Otherwise the output includes the gtable.
  
  # summary of the variable
  s <- summary(chemicals[[i]])
  
  # print markdown title: cat()
  cat(paste(chemical_names[[i]], '\n'))
  print(s)
}
```


## Bivariate Plots
```{r}
# 3 kinds of plot to explore 
# the relationship between a chemical and the quality

for (i in 1:11){
  
  # step 1: get arguments
  c1 <- chemicals[[i]]
  c2 <- chemicals_group_by_qty[[i]]
  
  c_name <- chemical_names[[i]]
  
  unit <- chemical_units[[i]]
  
  start <- chemical_args[[i]][[1]]
  end <- chemical_args[[i]][[2]]
  step <- chemical_args[[i]][[3]]
  
  title <- paste(c_name, "and Quality\n")
  
  # step 2: draw plots
  p1 <- create_scatter(c1, c_name, start, end, step, unit)
  p2 <- create_ridgeline(c2, c_name, start, end, step)
  p3 <- create_violin(c2, c_name, start, end, step)
  
  # step 3: print plots
  cat(title)
  print(p1)
  grid.arrange(p2, p3, nrow = 1)
}
```


### Spearman Correlation Coefficience: Quality vs Chemicals
```{r}
# Spearman Correlation: Quality vs Chemicals

# Build a new data frame to store r
cor_quality <- data.frame(matrix(nrow = 11, ncol = 2))

# Name 2 columns of table
colnames(cor_quality) <- c("chemicals", "r")

# Put r into the new data frame
for (i in 1:11) {
 cor <- cor(chemicals[[i]], df$quality, method = "spearman")
 cor_quality [i, 1] <- chemical_names[i]
 cor_quality [i, 2] <- cor
}
```

```{r}
# Show r in table

# Ordered by absolute value of r
cor_table <- arrange(cor_quality, desc(abs(r)))

# Print in table format
cor_table %>% kable() %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed", "responsive"), 
    full_width = F
  )

```

```{r}
# Plot of r

# Reorder
cor_quality %>%
  arrange(abs(r)) %>%
  #mutate(new_var = existing_var)
  mutate(c = factor(chemicals, chemicals)) %>%
  ggplot(aes(x = c, y = r)) +
    geom_segment( aes(x = c, xend=c, y=0, yend=r), 
                  color="skyblue", 
                  size=1) +
    geom_point( color="blue", size=4, alpha=0.6) +
    theme_light() +
    coord_flip() +
    theme(
      panel.grid.major.y = element_blank(),
      panel.border = element_blank(),
      axis.ticks.y = element_blank()
    ) +
  xlab("") +
  ylab("Spearman Correlation Coefficience with Quality")
```




## Multivariate Plots

### Correlation Matrix between Chemicals
```{r results = 'hide'}
# Compute 'Pearson' correlation between chemicals
df_no_quality <- df[, (2:12)]
r <- cor(df_no_quality)
head(round(r,2))
```

#### Correlation
```{r}
# Build a Pannel of 100 colors with Rcolor Brewer
r_colors <- brewer.pal(5, "RdBu")
r_colors=colorRampPalette(r_colors)(100)
 
# Order the correlation matrix
ord <- order(r[1, ])
r_ord = r[ord, ord]
plotcorr(r_ord , col=r_colors[r_ord*50+50] , mar=c(1,1,1,1)  )
```

#### Pearson Correlation Coefficient between Chemicals
```{r results='hide'}
#To compute the matrix of p-value, a custom R function is used :

# mat : is a matrix of data
# ... : further arguments to pass to the native R cor.test function
cor.mtest <- function(mat, ...) {
    mat <- as.matrix(mat)
    n <- ncol(mat)
    p.mat<- matrix(NA, n, n)
    diag(p.mat) <- 0
    for (i in 1:(n - 1)) {
        for (j in (i + 1):n) {
            tmp <- cor.test(mat[, i], mat[, j], ...)
            p.mat[i, j] <- p.mat[j, i] <- tmp$p.value
        }
    }
  colnames(p.mat) <- rownames(p.mat) <- colnames(mat)
  p.mat
}
# matrix of the p-value of the correlation
p.mat <- cor.mtest(df_no_quality)
head(p.mat[, 1:5])

### Source: http://www.sthda.com/english/wiki/visualize-correlation-matrix-using-correlogram
```

```{r}
col <- colorRampPalette(c("#BB4444", "#EE9988", "#FFFFFF", "#77AADD", "#4477AA"))

corrplot(r, method="color", col=col(1000),  
         type="upper", order="hclust", 
         addCoef.col = "black", # Add coefficient of correlation
         tl.col="black", tl.srt=20, # Text label color and rotation
         p.mat = p.mat, sig.level = 0.05, insig = "blank", # Combine with significance
         diag=FALSE,  # hide correlation coefficient on the principal diagonal
         number.cex=0.75, # change font size of r
         tl.cex = 0.8 # change font size of labels
         )

### source: http://www.sthda.com/english/wiki/visualize-correlation-matrix-using-correlogram
```

#### Moderately / Highly Correlated Chemicals
```{r}
# plot the chemicals with r>0.5
scatterplotMatrix(~citric.acid +
                    density +
                    fixed.acidity + 
                    free.sulfur.dioxide + 
                    total.sulfur.dioxide, 
                  data = df ,
                  reg.line= FALSE,
                  col= my_color , 
                  pch= 20,
                  cex=1.5,
                  main="Scatter Plot of Correlated Chemicals" )

```


### 

3 chemicals are correlated with quality:

+ Volatile Acidity: negative
+ Sulphates: positive
+ Alcohol: positive

```{r}
# Draw Scatter Plot: volatile.acidity & alcohol & quality
create_scatter_3var(df_alcohol$volatile.acidity, 'Volatile Acidity', 
                    chemical_units[[10]],
                    chemical_args[[10]][[1]],
                    chemical_args[[10]][[2]],
                    chemical_args[[10]][[3]],
                    df_alcohol$alcohol.group, 'Alcohol',
                    chemical_units[[1]])
```

```{r}
# Draw Scatter Plot: sulphates & alcohol & quality
create_scatter_3var(df_alcohol$sulphates, 'Sulphates', 
                    chemical_units[[9]],
                    chemical_args[[9]][[1]],
                    chemical_args[[9]][[2]],
                    chemical_args[[9]][[3]],
                    df_alcohol$alcohol.group, 'Alcohol',
                    chemical_units[[1]])
```